CF Registry Analysis

### 2024-11-12 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
model comparison table and plot

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(time = as.numeric(tstart), time2 = as.numeric(tstop), event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# Just split data by time period so we're only fitting 3 models
# "[2003,2009)", "[2009,2014)", "[2014,2019)"
annual$time_period <- cut(annual$ReviewYear,
  breaks = c(-Inf, 2009, 2014, Inf), right = F
)
era1 <- coxph(f, data = annual[as.numeric(annual$time_period) == 1, ], cluster = eDWID)
era2 <- coxph(f, data = annual[as.numeric(annual$time_period) == 2, ], cluster = eDWID)
era3 <- coxph(f, data = annual[as.numeric(annual$time_period) == 3, ], cluster = eDWID)
```

@@

```{r}
annual$tstart <- as.numeric(annual$tstart)
annual$tstop <- as.numeric(annual$tstop)
annual_split <- survSplit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ .,
  data = annual,
  cut = c(
    as.numeric(ymd("2008-12-31")) - as.numeric(og),
    as.numeric(ymd("2013-12-31")) - as.numeric(og)
  ),
  episode = "tgroup"
)
annual_split$tgroup <- factor(annual_split$tgroup,
  levels = 1:3, labels = c("[2003,2009)", "[2009,2014)", "[2014,2019)")
)
y <- "Surv(tstart, tstop, event)"
```

In the plots below, time-varying coefficients are excluded from the first forest plot, to show the average effect of the predictors. In the second forest plot, the estimated marginal means (EMMs) for the time-varying coefficients are shown. The EMMs are averaged over the levels of all the other variables in the model. In other words, the EMM represents the baseline hazard (not hazard ratio) at the average level of all the other variables in the model. For example, the log hazard for an average participant age [30, Inf) in the time period [2003,2009) is 1.887 and the log hazard for a similar participant in [2014,2019) is 1.401.

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression(exponentiate = T) %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split, new_f_emm, nuisance = xvars[-which(xvars == v)])
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_table(cox_res_split,
    show_p_values = F, add_reference_rows = T, exponentiate = T,
    signif_stars = F, colour = F
  ) %>%
    print()
  cat("\n")
  cat("\n")
}))
```

>>>>

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# Just split data by time period so we're only fitting 3 models
# "[2003,2009)", "[2009,2014)", "[2014,2019)"
annual$time_period <- cut(annual$ReviewYear,
  breaks = c(-Inf, 2009, 2014, Inf), right = F
)
era1 <- coxph(f,
  data = annual[as.numeric(annual$time_period) == 1, ],
  cluster = eDWID
)
era2 <- coxph(f,
  data = annual[as.numeric(annual$time_period) == 2, ],
  cluster = eDWID
)
era3 <- coxph(f,
  data = annual[as.numeric(annual$time_period) == 3, ],
  cluster = eDWID
)
# # KM plot
# fit <- survfit(Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
#   data = annual[as.numeric(annual$time_period) == 1, ]
# )
# # Visualize with survminer
# ggsurvplot(fit, data = annual[as.numeric(annual$time_period) == 1, ], risk.table = TRUE)
```

@@

#### 2003 - 2008

```{r}
t1 <- era1 %>% tbl_regression(exponentiate = T)
t1
ggcoef_model(era1,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

#### 2009 - 2013

```{r}
t2 <- era2 %>% tbl_regression(exponentiate = T)
t2
ggcoef_model(era2,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

#### 2014 - 2018

```{r}
t3 <- era3 %>% tbl_regression(exponentiate = T)
t3
ggcoef_model(era3,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

#### Combined results

```{r}
tbl_merge(
  tbls = list("2003 - 2008" = t1, "2009 - 2013" = t2, "2014 - 2018" = t3)
)
ggcoef_compare(
  list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
  add_reference_rows = T, exponentiate = T, type = "faceted", colour = F
)
```
